The Wolf in Sheep's Clothing:\\ Subversive Adversarial Agent Identification

### 2016-08-10 article.tex
Abstract and related work update

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Jean P. Castillo\IEEEauthorrefmark{1},
Matthew W. Swartwout\IEEEauthorrefmark{2}, Russel C. Jackson\IEEEauthorrefmark{3} and
Murat C. Cavusoglu\IEEEauthorrefmark{4}}
\IEEEauthorblockA{Department of Electrical Engineering and Computer Science,
Case Western Resere University\\
Cleveland, OH 44106\\
Email: \IEEEauthorrefmark{1}jpc87@case.edu,
\IEEEauthorrefmark{2}mws85@case.edu,
\IEEEauthorrefmark{3}rcj33@case.edu,
\IEEEauthorrefmark{4}mcc14@case.edu}}

@@

Mobile robot security is a rapidly developing area of interest in robotics. As mobile robotic systems become more present, the need for secure and reliable state estimation systems continues to grow. The increasing coverage of wireless networks and density of network cyber-physical systems has lead to an opportunity for distributed systems coordinate between multiple agents. This paper proposes a distributed state estimation system using Kalman filters with secure ad-hoc channel communications between the agents. On top of this is a state detection system validation allows agents to detect subversive or malfunctioning agents in their environment. The detection of such compromised agents can be used to create a \textit{trusted set} within a group, providing a multi-node trust system. A TurtleBot running Robot Operating System (ROS) platform is used for implementation and performance is also evaluated in the Gazebo simulator.

@@

\section{Related Work}
Chain of trusts, as pertaining to cyber systems, have been introduced in works such as Adi et al where a DNA clone resistant approach is taken throughout\cite{adi2009mechatronic}; a Physical Unclonable Function (PUF) was introduced which obtains physical differences between other electronic devices. Similar to this work the extended chain of trust this paper is the HSE which is used to verify other parts of the system much like Adi et al\cite{adi2009mechatronic}.

\subsection{State Estimation}
State estimation is a fundamental problem for robotics. Without knowing its location, a robot cannot be very useful. One common technique for robot localization is to use a Kalman Filter (KF). The KF is one of the most studied and most heavily utilized Bayesian filter \cite[39-81]{ProbabilisticRobotics}. At its core, a KF represents the state of a linear system as a multivariate normal distribution. By representing the state of the system as a normal distribution the filter can compactly represent the degree of belief of any possible state of the system.

The application of the KF to the task of robot localization has been studied intently \cite{Localization2003, Mohsin2014}. In Mohsin et al, it is explained to be so popular for three main reasons\cite{Mohsin2014}. First, it is optimal under certain reasonable and common assumptions, second that it is recursive which is memory efficient, and finally that it relies only on having knowledge of noisy sensor data and not being able to measure certain state variables directly.

For this paper, we utilize an Unscented Kalman Filter (UKF) for the task of sensor fusion and localization. The UKF provides better accuracy for non-linear systems than a KF or Extended Kalman Filter (EKF) and is less computationally complex \cite{Julier1997}.

REMOVED EXPANDED EKF/UKF DISCUSSION

@@

\subsubsection{Integrity}
The accuracy of a KF when the sensors are functioning properly is not in question, so we must look at when the sensors are not functioning properly. We propose two primary failure modes for data integrity: sensor failure and false data injection.

In a sensor failure scenario one or more of the sensors will stop providing all feedback. Luckily, this is something that KFs can handle without problem. The KF is always using all available information to generate its state estimate, and if one sensor fails the only result is that the accuracy of the filter will decrease. Obviously, if 100\% of sensors fail, the filter will become non-functional, but as long as at least one sensor remains operational the filter will not have difficulties. It is also worth nothing that, due to the design of the filter, the KF does not have issues with sensors going in and out of a failure state. When information is available it is used, and when it is not available the sensor is ignored.

Knowing that sensor failure is not a huge risk for the KF, sensor interference is a much more complex problem. If an attacker is somehow able to inject false data into the filter the pose estimate can be skewed. And this is a very real problem. The number of ways that false data could be added to a filter are too numerous to count, ranging from physically blocking a lidar sensor to hijacking the network packets of a wirelessly communicating sensor.

Mo et al and Yang et al both demonstrate that KFs are susceptible to false data injection attacks, and that even with failure detectors a clever adversary can craft their attack to bypass these detectors \cite{Mo2010}\cite{Yang2013}. Bezzo et al and Mo et al both attack this problem by adding extra steps into their KF algorithm\cite{Bezzo_2014}\cite{Mo2014}. Both successfully show that the filter output can be shielded from the effects of the attack.

@@

In the field of automobiles Lyu et al developed a protocol named PBA that is well guarded against high traffic and lossy wireless networks\cite{lyu2016pba}. The same work addresses DOS attack resistance and stores shortened MAC address signatures for memory consumption and rapid verification \cite{lyu2016pba}. Lyu et al takes a private public key pair approach which may be prone to further attacks as the base station could update their keys \cite{lyu2016pba}. Connectivity to the central station after the release is not taken to be necessary in this paper's approach.

\subsection{Distributed State Estimation}
As wireless networks become more and more omnipresent in our world, and given the increasing number and decreasing costs of mobile robots, the concept of a distributed state estimation system is very desirable. Given the rapid rise of these technologies though, the amount of research done into distributed systems is quite small compared to that for single robots. This was true years ago , and is still true today\cite{Parker2000}.

Traditionally, in order to increase the accuracy of state estimation, more sensors are added to a robot. However, increasing the number of sensors on a robot will increase the cost and complexity. And, where a robot operating by itself might need $n$ sensors for a sufficiently accurate state estimate, when operating in an environment surrounded by $m$ other robots that can share sensor data there is the possibility of needing far fewer sensors on each individual robot.

Most research into such distributed systems relies on the concept of cooperative positioning. That is, the robots coordinate their motions in order to use their sensors to determine the state of the other robots. Most often this involves one robot acting as a stationary landmark while the other robot moves. Kurazume et al's works all explore the idea of cooperative positioning\cite{Kurazume1994}\cite{Kurazume1996}\cite{Kurazume1998} \cite{Kurazume2000}. This is a valuable and effective concept when working with a team of coordinated robots. However many systems will not allow for such close cooperation. For example, autonomous cars on a highway will not be able to start and stop for each other and still get to their destination on time. And in rush hour in a city the processing power needed to coordinate a plan for the thousands of cars on the road would be incredible.

Other work into distributed systems still makes the assumption of a cooperative team. Sanderson et al and Roumeliotis et al both demonstrate a system where a KF is implemented that estimates the state of multiple robots in a system \cite{Sanderson1997}\cite{Roumeliotis2002}. However these states are computed relative to each other, considering the network of robots to be essentially one large robot with different parts. Continuing the highway example, cars will be moving in and out of communication with each other as some enter the highway and others exit. Thus the group of robots will be constantly changing and evolving, a challenge not tackled by these works.

The previously mentioned works are all examples of a decentralized, rather than distributed KF problem. Olfati-Saber et al makes clear the distinction between these two, saying \cite{Olfati-Saber2005}:
\begin{quote}"...[decentralized] methods require a complete network with all-to-all links. This solution is not scalable for large-scale sensor networks... Thus, decentralized Kalman filtering and distributed Kalman filtering are two separate problems. In the latter one, each node only is allowed to communicate with its nieghobrs on a graph $G$ that is connected but rather sparse."
\end{quote}
Olfati-Saber et al presents a solution to the distributed problem using consensus filters and multiple KFs\cite{Olfati-Saber2005}. However this work still requires adaptation of the KFs and extra computation, which is something we try to avoid.

>>>>

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Jean P. Castillo\IEEEauthorrefmark{1},
Matthew W. Swartwout\IEEEauthorrefmark{2}, Russell C. Jackson\IEEEauthorrefmark{3} and
Murat C. Cavusoglu\IEEEauthorrefmark{4}}
\IEEEauthorblockA{Department of Electrical Engineering and Computer Science,
Case Western Resere University\\
Cleveland, OH 44106\\
Email: \IEEEauthorrefmark{1}jpc87@case.edu,
\IEEEauthorrefmark{2}mws85@case.edu,
\IEEEauthorrefmark{3}rcj33@case.edu,
\IEEEauthorrefmark{4}mcc14@case.edu}}

@@

Mobile robot security is a rapidly developing area of interest in robotics. As mobile robotic systems become more ubiquitous, the need for security and reliability continues to grow. One avenue for advancement in this is distributed systems coordinated between multiple agents. This paper proposes a distributed state estimation system using Kalman filters with secure ad-hoc channel communications between the agents. On top of this is a state detection system validation allows agents to detect compromised agents in their environment. The detection of such compromised agents can be used to create a \textit{trusted set} within a group, providing a multi-node trust system. A TurtleBot running Robot Operating System (ROS) platform is used for implementation and performance is also evaluated in the Gazebo simulator.

@@

\section{Related Work}
Chain of trusts, as pertaining to cyber systems, have been introduced in works such as Adi et al where a DNA clone resistant approach is taken throughout~\cite{adi2009mechatronic}; a Physical Unclonable Function (PUF) was introduced which obtains physical differences between other electronic devices. Similar to this work the extended chain of trust this paper is the HSE which is used to verify other parts of the system much like Adi et al~\cite{adi2009mechatronic}.

\subsection{State Estimation}
State estimation is a fundamental problem for robotics. Without knowing its location, a robot cannot be very useful. One common technique for robot localization is to use a Kalman Filter (KF). The KF is one of the most studied and most heavily utilized Bayesian filter~\cite[39-81]{ProbabilisticRobotics}, and the application of the KF to the task of robot localization has been studied intently~\cite{Localization2003, Mohsin2014}. In \textcite{Mohsin2014}, it is explained to be so popular for three main reasons. First, it is optimal under certain reasonable and common assumptions, second that it is recursive which is memory efficient, and finally that it relies only on having knowledge of noisy sensor data and not being able to measure certain state variables directly~\cite{Mohsin2014}.

For this paper, we utilize an Unscented Kalman Filter (UKF) for the task of sensor fusion and localization. The UKF provides better accuracy for non-linear systems than a KF or Extended Kalman Filter (EKF) and is less computationally complex than the EKF~\cite{Julier1997}.

@@

\subsubsection{Integrity}
Integrity is the question of whether or not the state estimation system's output is accurate. We propose two primary failure modes for integrity: sensor failure and false data injection.

In a total sensor failure scenario one or more of the sensors will stop providing all feedback. This is something that KFs can handle without problem. The KF is always using all available information to generate its state estimate, and if one sensor fails the only result is that the accuracy of the filter will decrease. If 100\% of sensors fail, the filter will become non-functional, but this would affect any state estimation system. The KF also does not have issues with sensors going in and out of a failure state. When information is available it is used, and when it is not available the sensor can be ignored.

Sensor failure is not a huge risk for the KF, but  sensor interference is a much more complex problem. If an attacker is able to inject false data, the localization estimate can be skewed. This is a very real problem with many ways that false data could be added, ranging from physically blocking a LIDAR sensor to hijacking the network packets of a wireless sensor.

\textcite{Mo2010, Yang2013} demonstrate that KFs are susceptible to false data injection attacks, and that even with failure detectors, a clever adversary can craft their attack to bypass these detectors. \textcite{Bezzo_2014, Mo2014} attack this problem by adding extra steps to their KF algorithm. Both successfully show that the filter output can be shielded from the effects of the false data injection attack.

@@

In the field of automobiles Lyu et al developed a protocol named PBA that is well guarded against high traffic and lossy wireless networks~\cite{lyu2016pba}. The same work addresses DOS attack resistance and stores shortened MAC address signatures for memory consumption and rapid verification~\cite{lyu2016pba}. Lyu et al takes a private public key pair approach which may be prone to further attacks as the base station could update their keys~\cite{lyu2016pba}. Connectivity to the central station after the release is not taken to be necessary in this paper's approach.

\subsection{Distributed State Estimation}
As single robot systems coalesce into functional teams, distributed state estimation systems are desirable. Traditionally, in order to increase the accuracy of state estimation, more sensors are added to an individual robot. These additional sensors increase the cost and complexity of the robot. An alternative solution is for robots to share their sensor readings with others in their environment. This increases the amount of sensor data available to each robot, while keeping the number of sensors low on any one robot.

Most research into such systems relies on the concept of cooperative positioning~\cite{Kurazume1994} where robots coordinate their motions so that they can localize, or aid in the localization of, other robots. Often this involves one robot acting as a stationary landmark while the other robot moves. Studies done on this system have shown its effectiveness~\cite{Kurazume1996, Kurazume1998, Kurazume2000}. This is a valuable concept when working with a team of coordinated robots. However, many systems will not allow for cooperation.

Other work still makes the assumption of a cooperative team, even when they don't rely on cooperative positioning. \textcite{Sanderson1997, Roumeliotis2002} both demonstrate a system where a KF is implemented that estimates the state of multiple robots in a system. However these states are computed relative to each other, representing the network of robots as one large robot with different parts.

The previously mentioned works are all examples of a decentralized, rather than distributed KF problem. \textcite{Olfati-Saber2005} makes clear the distinction between these two, explaining that a decentralized method requires a complete connection between all nodes in the system, whereas a distributed system does not.

One truly distributed solution uses consensus filters and multiple KFs~\cite{Olfati-Saber2005}. However this still requires adaptation of the KFs and extra computation, which is something we try to avoid. In this paper we will show a distributed system with minimal KF modifications, and not a decentralized one.
